% Blueprint content: decomposed nodes (community-standard style).
%   \lean{decl}  links to a specific Lean declaration (verified to exist by checkdecls).
%   \leanok      used ONLY for nodes fully proved by the Lean kernel (no sorry) -> green in the graph.
%                Nodes with sorry / a library gap get only \lean (no \leanok) and state their status in prose.
%   \uses{...}   dependencies (edges of the dependency graph). \uses takes labels, so it is unaffected
%                by which environment (lemma/step/faithcheck) a node uses.
% ROLE-DISTINCT ENVIRONMENTS (see macros/common.tex):
%   lemma/theorem/definition = the PAPER's own statement (title cites the paper's number/section);
%   step                     = a sub-lemma decomposing a Lean proof (NOT a paper lemma);
%   faithcheck               = a machine-checked faithfulness audit (NOT paper math).
% The verbatim statement and proof are in docs/paper-lemma1.md (single source of truth);
% this chapter is its formalization decomposition. Language: English, matching the paper.

\chapter{Lemma 1: Posterior process}

\noindent\textbf{Legend --- what each node is.} This blueprint mixes three kinds of node; the
\emph{environment name} tells you which:
\begin{itemize}
  \item \textbf{Definition / Lemma / Theorem} --- a statement \emph{from the paper}. Its title cites
    the paper's own location or number. (E.g.\ the node ``Lemma~1'' below is the paper's Lemma~1.)
  \item \textbf{Step} --- a sub-lemma that only decomposes a Lean \emph{proof}; it is \emph{not} a
    lemma in the paper. Numbered on its own counter (Step~1, 2, \dots).
  \item \textbf{Faithfulness check} --- a machine-checked audit that the Lean \emph{encoding} matches
    the paper; it is not mathematical content of the paper. Own counter.
\end{itemize}
The actual formal object is always the \textbf{Lean declaration} reached via the
\textsf{L$\exists\forall$N} icon. The ``Lemma~$N$''/``Definition~$N$'' numbers here are internal to
this blueprint and are \emph{not} the paper's numbering, \emph{except} where a title says so.

\medskip
\noindent\fbox{\parbox{0.95\textwidth}{%
\textbf{What green does \emph{not} mean.} A green node means the Lean \textbf{kernel has checked the
proof against the stated Lean proposition} --- nothing more. It is \emph{not} a claim that the Lean
proposition faithfully says what the paper says. That second question (``statement $\leftrightarrow$
paper'') is a human translation step, and no colour on this page certifies it. What speaks to it is:
the bracketed tag on each node
(\emph{[faithful]}/\emph{[weakening]}/\emph{[strengthening]}/\emph{[encoding]}/\emph{[deviation]}),
the \textbf{Faithfulness check} nodes, and the verbatim paper text in \texttt{docs/paper-lemma1.md}
and \texttt{docs/paper-model.md}. Residual assumptions are listed explicitly rather than hidden:
see the tags marked \emph{[strengthening]} and \emph{[assumed]}.}}

\medskip
\noindent\textbf{Paper $\leftrightarrow$ Lean correspondence (top level).}
\begin{itemize}
  \item \textbf{Paper Lemma~1} (posterior process) $\;\leftrightarrow\;$ \texttt{kernel\_exists} $+$
    \texttt{factorization\_posterior} $+$ \texttt{testfun\_martingale}; its proof =
    Steps~\ref{lem:existence}, \ref{lem:sufficiency}, \ref{lem:factorization}, \ref{lem:martingale}.
  \item \textbf{Paper \S2.1 environment} $\leftrightarrow$ Definition~\ref{def:setup}.
    Belief space $\Delta(\Theta)\leftrightarrow$ \texttt{DMC.Belief}; allocation date
    $\leftrightarrow$ \texttt{DMC.AllocDate} (Model chapter).
  \item \textbf{Paper \S2.2 values} $U_t,V_t\leftrightarrow$ \texttt{DMC.Paper.U},
    \texttt{DMC.Paper.V} (the paper's defining suprema); $g\leftrightarrow$
    \texttt{DMC.DateValues.envelope}. \texttt{DMC.DateValues}/\texttt{DMC.SingleShot} are the
    downstream \emph{interface}, now derived via \texttt{DMC.Paper.toDateValues} (Model chapter).
  \item \emph{No paper counterpart:} every \textbf{Step} and \textbf{Faithfulness check} node --- these
    are formalization artifacts (proof decomposition, resp.\ encoding audits), not paper statements.
\end{itemize}

\noindent\textbf{How to read this page.} Hover over a statement's title bar to reveal a row of
small icons, then click the \textsf{L$\exists\forall$N} icon to open the Lean declaration that
formalizes it. A statement is coloured green exactly when the Lean kernel has fully verified it (no
\texttt{sorry}). \textbf{The verbatim statement and proof are in \texttt{docs/paper-lemma1.md}};
this chapter decomposes that proof, following the paper's own four steps.

\begin{definition}[Environment (paper \S2.1)]\label{def:setup}
    Discrete time $t\in\mathbb T=\{0,1,2,\dots\}$. A Borel probability space
    $(\Omega,\mathcal F,\mathbb P)$; primitive state $\theta:\Omega\to\Theta$ with $\Theta$
    compact Polish; $\Delta(\Theta)$ the space of Borel probability measures on $\Theta$ (weak
    topology). Public state $Z_t:\Omega\to E_t$ ($E_t$ Polish), with $\mathcal F_t=\sigma(Z_t)$
    an increasing filtration and $\mathcal F_0$ trivial.
\end{definition}

%======================================================================
% Main statement (paper Lemma 1, verbatim); its proof is decomposed into the four step-nodes below.
%======================================================================
\begin{lemma}[Posterior process --- \textbf{paper Lemma 1}]\label{lem:posterior-martingale}
    \lean{PosteriorProcess.kernel_exists, PosteriorProcess.factorization_posterior,
      PosteriorProcess.testfun_martingale}\leanok
    \uses{def:setup}
    For each $t\in\mathbb T$, there exists an $\mathcal F_t$-measurable posterior kernel
    $S_t:\Omega\to\Delta(\Theta)$ such that, for every bounded Borel function
    $\varphi:\Theta\to\mathbb R$,
    \[
      \int_\Theta \varphi(\theta')\,S_t(\omega)(d\theta')
        = \mathbb E[\varphi(\theta)\mid\mathcal F_t](\omega)\qquad\text{a.s.}
    \]
    In particular:
    \textbf{(i)} $S_t$ is a sufficient statistic for $\theta$ given $\mathcal F_t$;
    \textbf{(ii)} since $\mathcal F_t = \sigma(Z_t)$, the Doob--Dynkin lemma implies that
    $S_t = f_t(Z_t)$ for some Borel measurable $f_t$;
    \textbf{(iii)} $\{S_t\}_{t\in\mathbb T}$ is a bounded $(\mathcal F_t)$-martingale.

% BEGIN-LEAN-STMT (auto-generated by tools/gen_lean_statements.py -- do not edit)
\par\smallskip\noindent\textbf{Lean statements (printed by Lean; verbatim):}
\begin{verbatim}
@PosteriorProcess.kernel_exists : ∀ {Ω : Type u_1} {m : MeasurableSpace Ω} {P : MeasureTheory.Measure Ω}
  [MeasureTheory.IsProbabilityMeasure P] {ι : Type u_2} [inst : Preorder ι] (ℱ : MeasureTheory.Filtration ι m)
  {Θ : Type u_3} [inst_1 : MeasurableSpace Θ] [StandardBorelSpace Θ] [Nonempty Θ] (θ : Ω → Θ),
  Measurable θ →
    ∀ (t : ι),
      ∃ S,
        (∀ (ω : Ω), MeasureTheory.IsProbabilityMeasure (S ω)) ∧
          Measurable S ∧
            ∀ (φ : Θ → ℝ),
              Measurable φ → (∃ C, ∀ (x : Θ), |φ x| ≤ C) → (fun ω => ∫ (x : Θ), φ x ∂S ω) =ᵐ[P] P[φ ∘ θ | ↑ℱ t]
\end{verbatim}
\begin{verbatim}
@PosteriorProcess.factorization_posterior : ∀ {Ω : Type u_1} {E : Type u_2} {Θ : Type u_3} [inst : MeasurableSpace E]
  [inst_1 : TopologicalSpace Θ] [PolishSpace Θ] [CompactSpace Θ] [inst_4 : MeasurableSpace Θ] [BorelSpace Θ]
  [Nonempty (MeasureTheory.ProbabilityMeasure Θ)] (Z : Ω → E) (S : Ω → MeasureTheory.ProbabilityMeasure Θ),
  Measurable S → ∃ f, Measurable f ∧ S = f ∘ Z
\end{verbatim}
\begin{verbatim}
@PosteriorProcess.testfun_martingale : ∀ {Ω : Type u_1} {m : MeasurableSpace Ω} {P : MeasureTheory.Measure Ω}
  [MeasureTheory.IsProbabilityMeasure P] {ι : Type u_2} [inst : Preorder ι] (ℱ : MeasureTheory.Filtration ι m)
  {Θ : Type u_3} [inst_1 : MeasurableSpace Θ] (θ : Ω → Θ) (φ : Θ → ℝ),
  Measurable φ → (∃ C, ∀ (x : Θ), |φ x| ≤ C) → MeasureTheory.Martingale (fun t => P[φ ∘ θ | ↑ℱ t]) ℱ P
\end{verbatim}
% END-LEAN-STMT
\end{lemma}
\begin{proof}\leanok
    \uses{lem:existence, lem:sufficiency, lem:factorization, lem:martingale}
    The statement is the conjunction of the four steps below: existence and the kernel property
    (Step~\ref{lem:existence}), sufficiency (Step~\ref{lem:sufficiency}, part~(i)),
    factorization through $Z_t$ (Step~\ref{lem:factorization}, part~(ii)), and the martingale
    property (Step~\ref{lem:martingale}, part~(iii)). The verbatim proof is in
    \texttt{docs/paper-lemma1.md}.
\end{proof}

%======================================================================
% Step 1: existence + kernel property
%======================================================================
\begin{step}[Existence of the posterior kernel]\label{lem:existence}
    \lean{PosteriorProcess.kernel_exists}\leanok
    \uses{def:setup}
    For each $t$, there exists an $\mathcal F_t$-measurable $S_t:\Omega\to\Delta(\Theta)$ such
    that, for every bounded Borel $\varphi$, the map
    $\omega\mapsto\int_\Theta\varphi\,dS_t(\omega)$ equals
    $\mathbb E[\varphi(\theta)\mid\mathcal F_t]$ almost surely.

% BEGIN-LEAN-STMT (auto-generated by tools/gen_lean_statements.py -- do not edit)
\par\smallskip\noindent\textbf{Lean statement (printed by Lean; verbatim):}
\begin{verbatim}
@PosteriorProcess.kernel_exists : ∀ {Ω : Type u_1} {m : MeasurableSpace Ω} {P : MeasureTheory.Measure Ω}
  [MeasureTheory.IsProbabilityMeasure P] {ι : Type u_2} [inst : Preorder ι] (ℱ : MeasureTheory.Filtration ι m)
  {Θ : Type u_3} [inst_1 : MeasurableSpace Θ] [StandardBorelSpace Θ] [Nonempty Θ] (θ : Ω → Θ),
  Measurable θ →
    ∀ (t : ι),
      ∃ S,
        (∀ (ω : Ω), MeasureTheory.IsProbabilityMeasure (S ω)) ∧
          Measurable S ∧
            ∀ (φ : Θ → ℝ),
              Measurable φ → (∃ C, ∀ (x : Θ), |φ x| ≤ C) → (fun ω => ∫ (x : Θ), φ x ∂S ω) =ᵐ[P] P[φ ∘ θ | ↑ℱ t]
\end{verbatim}
% END-LEAN-STMT
\end{step}
\begin{proof}\leanok
    $\Theta$ compact Polish $\Rightarrow$ standard Borel; the claim follows from the existence
    of regular conditional distributions (Kallenberg 2002, Thm~6.3). In Lean the kernel is the
    regular conditional distribution \texttt{condDistrib $\theta$ id P} of $\theta$ given
    $\mathcal F_t$. Verified by the Lean kernel, no \texttt{sorry}.
\end{proof}

%======================================================================
% Step 2: (i) sufficiency
%======================================================================
\begin{step}[Part (i): Sufficiency]\label{lem:sufficiency}
    \lean{PosteriorProcess.sufficiency}\leanok
    \uses{lem:existence}
    Conditioning on $S_t$ and conditioning on $\mathcal F_t$ yield the same conditional
    distribution of $\theta$; that is, $S_t$ is a sufficient statistic for $\theta$ given
    $\mathcal F_t$.

% BEGIN-LEAN-STMT (auto-generated by tools/gen_lean_statements.py -- do not edit)
\par\smallskip\noindent\textbf{Lean statement (printed by Lean; verbatim):}
\begin{verbatim}
@PosteriorProcess.sufficiency : ∀ {Ω : Type u_1} {m : MeasurableSpace Ω} {P : MeasureTheory.Measure Ω}
  [MeasureTheory.IsProbabilityMeasure P] {ι : Type u_2} [inst : Preorder ι] (ℱ : MeasureTheory.Filtration ι m)
  {Θ : Type u_3} [inst_1 : MeasurableSpace Θ] [StandardBorelSpace Θ] [Nonempty Θ] (θ : Ω → Θ),
  Measurable θ →
    ∀ (t : ι),
      ∃ S,
        (∀ (ω : Ω), MeasureTheory.IsProbabilityMeasure (S ω)) ∧
          Measurable S ∧
            ∀ (φ : Θ → ℝ),
              Measurable φ →
                (∃ C, ∀ (x : Θ), |φ x| ≤ C) → P[φ ∘ θ | ↑ℱ t] =ᵐ[P] P[φ ∘ θ | MeasurableSpace.comap S inferInstance]
\end{verbatim}
% END-LEAN-STMT
\end{step}
\begin{proof}\leanok
    For bounded Borel $\varphi$, the map $F(\mu):=\int\varphi\,d\mu$ is Borel measurable on
    $\Delta(\Theta)$ by the functional monotone class theorem (it holds for continuous $\varphi$
    by the definition of the weak topology, and extends to bounded Borel $\varphi$ by dominated
    convergence). Hence $\mathbb E[\varphi(\theta)\mid\mathcal F_t]=F(S_t)$ is
    $\sigma(S_t)$-measurable, and uniqueness of conditional expectation gives the equality of the
    two conditional expectations. Verified by the Lean kernel, no \texttt{sorry}.
\end{proof}

%======================================================================
% Step 3: (ii) Doob-Dynkin -- abstract layer (proved) + library gap + full version
%======================================================================
\begin{step}[Abstract Doob--Dynkin factorization]\label{lem:doob-abstract}
    \lean{PosteriorProcess.factorization}\leanok
    If $S$ is $\sigma(Z)$-measurable and takes values in a standard Borel space, then
    $S=f\circ Z$ for some Borel measurable $f$.

% BEGIN-LEAN-STMT (auto-generated by tools/gen_lean_statements.py -- do not edit)
\par\smallskip\noindent\textbf{Lean statement (printed by Lean; verbatim):}
\begin{verbatim}
@PosteriorProcess.factorization : ∀ {Ω : Type u_1} {E : Type u_2} {Δ : Type u_3} [inst : MeasurableSpace E]
  [inst_1 : MeasurableSpace Δ] [StandardBorelSpace Δ] [Nonempty Δ] (Z : Ω → E) (S : Ω → Δ),
  Measurable S → ∃ f, Measurable f ∧ S = f ∘ Z
\end{verbatim}
% END-LEAN-STMT
\end{step}
\begin{proof}\leanok
    Mathlib's \texttt{Measurable.exists\_eq\_measurable\_comp}. Verified by the Lean kernel, no
    \texttt{sorry}.
\end{proof}

\begin{step}[$\Delta(\Theta)$ is standard Borel]\label{lem:deltaTheta-sb}
    \lean{PosteriorProcess.deltaTheta_standardBorel}\leanok
    \uses{def:setup}
    $\Delta(\Theta)$ (Borel probability measures on the compact Polish space $\Theta$, weak
    topology) is a standard Borel space.

% BEGIN-LEAN-STMT (auto-generated by tools/gen_lean_statements.py -- do not edit)
\par\smallskip\noindent\textbf{Lean statement (printed by Lean; verbatim):}
\begin{verbatim}
PosteriorProcess.deltaTheta_standardBorel : ∀ (Θ : Type u_1) [inst : TopologicalSpace Θ] [PolishSpace Θ]
  [CompactSpace Θ] [inst_3 : MeasurableSpace Θ] [BorelSpace Θ], StandardBorelSpace (MeasureTheory.ProbabilityMeasure Θ)
\end{verbatim}
% END-LEAN-STMT
\end{step}
\begin{proof}\leanok
    This was a gap in Mathlib, now proved from scratch in two halves: $\Delta(\Theta)$ carries a
    Polish topology (compact by Prokhorov + metrizable by Lévy--Prokhorov, hence completely
    metrizable), and its Giry $\sigma$-algebra coincides with the Borel $\sigma$-algebra of the
    weak topology. Verified by the Lean kernel, no \texttt{sorry}.
\end{proof}

\begin{faithcheck}[$\Delta(\Theta)$ is the paper's object: its σ-algebra and its topology]
    \label{fc:deltaTheta-object}
    \lean{PosteriorProcess.deltaTheta_measurableSpace_eq_borel,
      PosteriorProcess.deltaTheta_tendsto_iff_integral}\leanok
    \uses{lem:deltaTheta-sb}
    Step~\ref{lem:deltaTheta-sb} is stated about Mathlib's \texttt{ProbabilityMeasure}~$\Theta$,
    which carries Mathlib's \emph{Giry} measurable structure and Mathlib's topology of convergence
    in distribution. The paper's $\Delta(\Theta)$ is the probability measures \emph{with the weak
    topology} (and, for the Doob--Dynkin step, that topology's Borel σ-algebra). A priori these are
    different structures, so a theorem about the wrong one would be true and useless. Two checks
    close the gap:
    \begin{itemize}
      \item \textbf{σ-algebra}: the measurable structure on $\Delta(\Theta)$ \emph{equals} the Borel
        σ-algebra of the weak topology --- the content of \texttt{deltaTheta\_borelSpace}, spelled
        out as an equation rather than hidden in a \texttt{BorelSpace} instance.
      \item \textbf{Topology}: convergence in $\Delta(\Theta)$ holds exactly when
        $\int f\,d\mu$ converges for every bounded continuous $f$ --- the defining property of the
        weak topology.
    \end{itemize}
    These are not extra assumptions; Mathlib's structures already are the right ones. The point is
    that this is now kernel-checked and visible, rather than silently trusted.

% BEGIN-LEAN-STMT (auto-generated by tools/gen_lean_statements.py -- do not edit)
\par\smallskip\noindent\textbf{Lean statements (printed by Lean; verbatim):}
\begin{verbatim}
PosteriorProcess.deltaTheta_measurableSpace_eq_borel : ∀ (Θ : Type u_1) [inst : TopologicalSpace Θ] [PolishSpace Θ]
  [CompactSpace Θ] [inst_3 : MeasurableSpace Θ] [inst_4 : BorelSpace Θ],
  inferInstance = borel (MeasureTheory.ProbabilityMeasure Θ)
\end{verbatim}
\begin{verbatim}
@PosteriorProcess.deltaTheta_tendsto_iff_integral : ∀ {Θ : Type u_1} [inst : TopologicalSpace Θ]
  [inst_1 : MeasurableSpace Θ] [inst_2 : OpensMeasurableSpace Θ] {γ : Type u_2} {F : Filter γ}
  {μs : γ → MeasureTheory.ProbabilityMeasure Θ} {μ : MeasureTheory.ProbabilityMeasure Θ},
  Filter.Tendsto μs F (nhds μ) ↔
    ∀ (f : BoundedContinuousFunction Θ ℝ),
      Filter.Tendsto (fun i => ∫ (x : Θ), f x ∂↑(μs i)) F (nhds (∫ (x : Θ), f x ∂↑μ))
\end{verbatim}
% END-LEAN-STMT
\end{faithcheck}

\begin{step}[Part (ii): Induction by $Z_t$]\label{lem:factorization}
    \lean{PosteriorProcess.factorization_posterior}\leanok
    \uses{lem:existence, lem:doob-abstract, lem:deltaTheta-sb}
    Since $\mathcal F_t=\sigma(Z_t)$, there is a Borel measurable $f_t:E_t\to\Delta(\Theta)$ with
    $S_t=f_t(Z_t)$ a.s.; the public history $Z_t$ induces the posterior $S_t$.

% BEGIN-LEAN-STMT (auto-generated by tools/gen_lean_statements.py -- do not edit)
\par\smallskip\noindent\textbf{Lean statement (printed by Lean; verbatim):}
\begin{verbatim}
@PosteriorProcess.factorization_posterior : ∀ {Ω : Type u_1} {E : Type u_2} {Θ : Type u_3} [inst : MeasurableSpace E]
  [inst_1 : TopologicalSpace Θ] [PolishSpace Θ] [CompactSpace Θ] [inst_4 : MeasurableSpace Θ] [BorelSpace Θ]
  [Nonempty (MeasureTheory.ProbabilityMeasure Θ)] (Z : Ω → E) (S : Ω → MeasureTheory.ProbabilityMeasure Θ),
  Measurable S → ∃ f, Measurable f ∧ S = f ∘ Z
\end{verbatim}
% END-LEAN-STMT
\end{step}
\begin{proof}\leanok
    Apply the abstract factorization (Step~\ref{lem:doob-abstract}) to $\Delta(\Theta)$, which
    is standard Borel (Step~\ref{lem:deltaTheta-sb}). Verified by the Lean kernel, no
    \texttt{sorry}.
\end{proof}

%======================================================================
% Step 4: (iii) martingale property (proved)
%======================================================================
\begin{step}[Part (iii): Martingale property]\label{lem:martingale}
    \lean{PosteriorProcess.testfun_martingale}\leanok
    \uses{lem:existence}
    For every bounded Borel $\varphi$, the real-valued process
    $t\mapsto\int_\Theta\varphi\,dS_t$ is an $(\mathcal F_t)$-martingale; hence $\{S_t\}$ is a
    bounded measure-valued martingale.

% BEGIN-LEAN-STMT (auto-generated by tools/gen_lean_statements.py -- do not edit)
\par\smallskip\noindent\textbf{Lean statement (printed by Lean; verbatim):}
\begin{verbatim}
@PosteriorProcess.testfun_martingale : ∀ {Ω : Type u_1} {m : MeasurableSpace Ω} {P : MeasureTheory.Measure Ω}
  [MeasureTheory.IsProbabilityMeasure P] {ι : Type u_2} [inst : Preorder ι] (ℱ : MeasureTheory.Filtration ι m)
  {Θ : Type u_3} [inst_1 : MeasurableSpace Θ] (θ : Ω → Θ) (φ : Θ → ℝ),
  Measurable φ → (∃ C, ∀ (x : Θ), |φ x| ≤ C) → MeasureTheory.Martingale (fun t => P[φ ∘ θ | ↑ℱ t]) ℱ P
\end{verbatim}
% END-LEAN-STMT
\end{step}
\begin{proof}\leanok
    By the tower property, $\mathbb E[\int\varphi\,dS_t\mid\mathcal F_s]
    =\mathbb E[\varphi(\theta)\mid\mathcal F_s]=\int\varphi\,dS_s$ for $s\le t$. Mathlib's
    \texttt{martingale\_condExp}. Verified by the Lean kernel, no \texttt{sorry}.
\end{proof}
